---
title: "ETC genes network plot DEG 17may R Notebook"
output: html_notebook
---
# First step: collected gene list from https://www.wikipathways.org/index.php/Pathway:WP111 in our dataset. The figure is inspired from https://www.pnas.org/content/114/32/8596/tab-figures-data (Fig 2C).
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("STRINGdb")
```
# For LDA, FOLLOWED:https://rpubs.com/ifn1411/LDA
```{r}
install.packages("MASS")
install.packages("car")
install.packages("caret")
install.packages("klaR")
install.packages("rattle.data")

```

```{r}
library(MASS)
library(car)
```

```{r}
setwd("D:/OneDrive - University of Pittsburgh/Amrita_revision_nature_aging/nature_aging_reviewer_comments_april2021")

library(STRINGdb)
library(readxl)
library(tidyverse)
library(igraph)
library("DESeq2")

string_db <- STRINGdb$new( version="11", species=10090, score_threshold=200, input_directory="")

```
# Perform deseq2 on dataset excluding outliers
```{r}
cts <- as.matrix(read.csv("unnormalized_counts_Amrita.txt",sep="\t"))

cts_new <- subset(cts, select = -c(OYS5, ODS5) )


coldata <-  as.data.frame(read.csv("annotation_file_wo_outliers.txt",sep="\t" ,row.names = 1))
#coldata <- coldata[,c("condition")]

coldata$condition <- factor(coldata$condition)

all(rownames(coldata) == colnames(cts_new)) #should get true


dds_new <- DESeqDataSetFromMatrix(countData = round(cts_new),
                              colData = coldata,
                              design = ~ condition)
dds_new

# Set reference control group
dds_new$condition <- relevel(dds_new$condition, ref = "Sham")
nrow(dds_new)


# Pre-filtering
keep <- rowSums(counts(dds_new)) >= 10
dds_new <- dds_new[keep,]
nrow(dds_new)

dds_new <- DESeq(dds_new)
res_new <- results(dds_new)
res_new
```
```{r}
DGE_results_DSvsSH <- results(dds_new, alpha = 0.1, contrast = c("condition", "DS", "SH" ) )
DGE_results_DSvsSH <- na.omit(DGE_results_DSvsSH)

DGE_results_YSvsSH <- results(dds_new, alpha = 0.1, contrast = c("condition", "YS", "SH" ) )
DGE_results_YSvsSH <- na.omit(DGE_results_YSvsSH)

summary(DGE_results_DSvsSH)

summary(DGE_results_YSvsSH)

YSvsDS_0.1 <- as.data.frame(DGE_results_YSvsDS_lfc_one)

YSvsDS_0.1change <- YSvsDS_0.1 %>%   filter(padj<0.1) 

DSvsSH_0.1 <- as.data.frame(DGE_results_DSvsSH_lfc_one)

DSvsSH_0.1up <- DSvsSH_0.1 %>%  filter(log2FoldChange>0.1)%>%  filter(padj<0.1)
DSvsSH_0.1down <- DSvsSH_0.1 %>%  filter(log2FoldChange< -0.1)%>%  filter(padj<0.1)

YSvsSH_0.1 <- as.data.frame(DGE_results_YSvsSH_lfc_one)

YSvsSH_0.1up <- YSvsSH_0.1 %>%  filter(log2FoldChange>0.1)%>%  filter(padj<0.1)
YSvsSH_0.1down <- YSvsSH_0.1 %>%  filter(log2FoldChange< -0.1)%>%  filter(padj<0.1)

onlyYSup<- setdiff(rownames(YSvsSH_0.1up),union(rownames(DSvsSH_0.1up),rownames(DSvsSH_0.1down)))
onlyYSdown<- setdiff(rownames(YSvsSH_0.1down),union(rownames(DSvsSH_0.1up),rownames(DSvsSH_0.1down)))

onlyYSupanddown <- setdiff(union(rownames(YSvsSH_0.1up),rownames(YSvsSH_0.1down)),union(rownames(DSvsSH_0.1up),rownames(DSvsSH_0.1down)))
```



# get complex1 genes - intersection with DE genes YS vs Sham
# first run deseq2 after removing outliers 
```{r}
complexes_cpm <-read_excel('meetings and data/ETC_complexes_genelist_17may.xlsx')

comp1_genes <- as_tibble(complexes_cpm) %>%  filter(Complex == 1) %>% dplyr::select(Gene)  

comp1_deg <- intersect(comp1_genes$Gene, onlyYSupanddown)
```
```{r}
comp2_genes <- as_tibble(complexes_cpm) %>%  filter(Complex == 2) %>% dplyr::select(Gene)  

comp2_deg <- intersect(comp2_genes$Gene, onlyYSupanddown)
```
```{r}
comp3_genes <- as_tibble(complexes_cpm) %>%  filter(Complex == 3) %>% dplyr::select(Gene)  

comp3_deg <- intersect(comp3_genes$Gene, onlyYSupanddown)
```
```{r}
comp4_genes <- as_tibble(complexes_cpm) %>%  filter(Complex == 4) %>% dplyr::select(Gene)  

comp4_deg <- intersect(comp4_genes$Gene, onlyYSupanddown)
```
```{r}
comp5_genes <- as_tibble(complexes_cpm) %>%  filter(Complex == 5) %>% dplyr::select(Gene)  

comp5_deg <- intersect(comp5_genes$Gene, onlyYSupanddown)
```

```{r}
comp1_net<-read.csv('ETC_network/comp1_stringdb.tsv',sep='\t', header = TRUE)

comp1_links <- data.frame(source=comp1_net$X.node1,target=comp1_net$node2)

comp1_network <- graph_from_data_frame(d=comp1_links, directed=F) 
plot(comp1_network, vertex.color="white",vertex.label.color='black')
```
```{r}
comp2_net<-read.csv('ETC_network/comp2_stringdb.tsv',sep='\t', header = TRUE)

comp2_links <- data.frame(source=comp2_net$X.node1,target=comp2_net$node2)

comp2_network <- graph_from_data_frame(d=comp2_links, directed=F) 
plot(comp2_network, vertex.color="white",vertex.label.color='black')
```
```{r}
comp3_net<-read.csv('ETC_network/comp3_stringdb.tsv',sep='\t', header = TRUE)

comp3_links <- data.frame(source=comp3_net$X.node1,target=comp3_net$node2)

comp3_network <- graph_from_data_frame(d=comp3_links, directed=F) 
plot(comp3_network, vertex.color="white",vertex.label.color='black')
```
```{r}
comp4_net<-read.csv('ETC_network/comp4_stringdb.tsv',sep='\t', header = TRUE)

comp4_links <- data.frame(source=comp4_net$X.node1,target=comp4_net$node2)

comp4_network <- graph_from_data_frame(d=comp4_links, directed=F) 
V(comp4_network)$color <-c('white')
V(comp4_network)["Cox6a1"]$color <- "red"


plot(comp4_network,vertex.label.color='black')

```
```{r}
comp5_net<-read.csv('ETC_network/comp5_stringdb.tsv',sep='\t', header = TRUE)

comp5_links <- data.frame(source=comp4_net$X.node1,target=comp4_net$node2)

comp5_network <- graph_from_data_frame(d=comp5_links, directed=F) 
plot(comp5_network, vertex.color="white",vertex.label.color='black')
```

# Representing LFC as color of nodes with only FDR<0.1 cutoff
```{r}
DGE_results_YSvsSHdf<-rownames_to_column(as.data.frame(DGE_results_YSvsSH), var = "rowname")
DGE_results_DSvsSHdf<-rownames_to_column(as.data.frame(DGE_results_DSvsSH), var = "rowname")

YSvsSH_0.1change <- as_tibble(DGE_results_YSvsSHdf) %>%   filter(padj<0.1)
DSvsSH_0.1change <- as_tibble(DGE_results_DSvsSHdf) %>%   filter(padj<0.1)

YSvsSH_genes <- (YSvsSH_0.1change) %>% dplyr::select(rowname)

DSvsSH_genes <- (DSvsSH_0.1change) %>% dplyr::select(rowname)



onlyYS_genes <- setdiff(YSvsSH_genes$rowname,DSvsSH_genes$rowname)

comp1_ys <- intersect(comp1_genes$Gene, onlyYS_genes)
comp2_ys <- intersect(comp2_genes$Gene, onlyYS_genes)
comp3_ys <- intersect(comp3_genes$Gene, onlyYS_genes)
comp4_ys <- intersect(comp4_genes$Gene, onlyYS_genes)
comp5_ys <- intersect(comp5_genes$Gene, onlyYS_genes)

```

# Fatty acid oxidation - gene network
```{r}
fattyacid_df <- read_excel('meetings and data/fatty_acid_beta_oxidation_73genes.xlsx')

fattyacid_allgenes <- fattyacid_df %>% dplyr::select(Symbol)

fattyacid_go <- enrichGO(gene= fattyacid_allgenes$Symbol,
               keyType       = 'SYMBOL',
               OrgDb         = org.Mm.eg.db,
               ont           = "BP",
#               readable      = TRUE,
               pAdjustMethod = "BH"
               )

dotplot(fattyacid_go) + ggtitle("verifying GO Terms")
```

#LDA data preprocessing
# followed this:https://rpubs.com/ifn1411/LDA
```{r}

# Pre-filtering
keep <- rowSums(counts(dds_new)) >= 10
dds_new2 <- dds_new[keep,]
nrow(dds_new2)

esf_dds2 <- estimateSizeFactors(dds_new2)

rlog_dds <- rlog(esf_dds2, blind = TRUE)

pc_serum <- prcomp(t(assay(rlog_dds)), scale=TRUE, center =TRUE)

pc_serum_6 <- (pc_serum$x[,1:6])

pc_serum_vars<- as.data.frame(pc_serum_6)

pc_serum_vars$condition <- factor(coldata$condition)

summary(pc_serum)

head(pc_serum_vars)

# lda_df <- data.frame(t(assay(rlog_dds)))
# lda_df$condition <- factor(coldata$condition)

# options(expressions = 5e5)
# newlda_df <- lda_df %>%
#   dplyr::select(condition, everything())
lda_df.lda <- lda(condition ~., data = pc_serum_vars)
lda_df.lda
```
```{r}
df.lda.values <- predict(lda_df.lda)
df.lda.values

plot(df.lda.values$x[,1], df.lda.values$x[,2])
```
```{r}
df.lda.values <- predict(lda_df.lda, dimen=3, method="predictive")

df.lda.values
plot(df.lda.values$x[,1], df.lda.values$x[,2])
```
```{r}
install.packages("ggalt")
library(ggalt)
```

```{r}
#convert to data frame 
newdata <- data.frame(condition = factor(coldata$condition), lda = df.lda.values$x)
library(ggplot2)



tiff('LDA_2d_encircle.tiff', units="in", width=6, height=4, res=500, compression = 'lzw')
ldaplot<- ggplot(newdata,aes(x=lda.LD1, y=lda.LD2, color = condition), size = 2.5) + geom_point(aes(x=lda.LD1, y=lda.LD2, color = condition), size = 4) +
  geom_encircle(aes(group = condition))+
  ylim(-4, 3)+xlim(-5,5)
print(ldaplot)
dev.off()

```


